---
title: "RPolarAverage3"
output: html_document
date: "2023-11-01"
---

```{r}
rm(list=ls())
ls()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(RColorBrewer)
knitr::opts_chunk$set(echo = TRUE)
#library(readtextgrid)
library(devtools)
load_all("/Users/yayaorigami/Documents/readtextgrid")
library(tibble)
library(dplyr)
#library(jsonlite)
library(rjson)
library(eply)
#library(textgRid)
library(extrafont)
library(Cairo)
#font_import()
# Load the DejaVu Sans font family
#loadfonts()

```

```{r}
loadTraces <- function(directory_name, tiername = "", categories = c()){

  metaDataFile <- paste(directory_name, "metadata.json", sep = "/") #is this an ok approach?
  #is it guaranteed that the meta file will be called metadata, and is in the same directory level

  if (!file.exists(metaDataFile)){
    stop("metadata file does not exist in given directory")
  }

  metaData <- fromJSON(file = metaDataFile)

  traces_raw <- metaData$traces["Tongue"]
  filesAll <- metaData$files

  myXY_data <- data.frame()
  column_names <- c('file_number', 'itemNumber_inFile', 'segment', 'x', 'y')
  myXY_data <- rbind(myXY_data, column_names)
  allRowsTextGrids <- list()

  for (individualFile in 1:length(filesAll)){
    #accessing text grid files
    recording_name <- filesAll[[individualFile]]$.TextGrid
    plainTextname <- filesAll[[individualFile]]$name
    #used to access text grid files
    fullFilePath <- file.path(directory_name, recording_name)
    #print(fullFilePath)

    cur_recording <- metaData$traces$tongue$files[[(filesAll[[individualFile]]$name)]]

    if (is.null(cur_recording)){
      listExists <- 0
    }else{
      listExists <- max(unlist(lapply(metaData$traces$tongue$files[[(filesAll[[individualFile]]$name)]], length)))
    }

    #read text grid if it is not empty
    if (listExists > 1){
         #read in text grid data but only if recording > 0
      #Reading Text Grid
      errorCode = 0
      tryCatch({
        textGridDataFile <- read_textgrid(fullFilePath)
      }, error = function(e){
        print("cannot open textgrid")
        errorCode = 1
      })
      if(errorCode==1){
        next
      }
      
      #time to parse
      intervalData <- textGridDataFile[textGridDataFile$tier_type == "IntervalTier", ]
      if (nchar(tiername)!=0){
        if (tiername %in% intervalData$tier_name){
          intervalData <- intervalData[intervalData$tier_name == tiername,]
        }else{ #you have annotation, but not in the proper tier
          print("skipping this textgrid file, as annotations are not in the proper tier")
          print(plainTextname)
          next
        }
      }
      # if (nchar(tiername)!=0){
      #   intervalData <- intervalData[intervalData$tier_name == tiername,]
      # }

      if (length(categories) == 0){ #read all segments
        #intervalData <- intervalData[nchar(intervalData$text) == 1 | nchar(intervalData$text) == 2, ] #just gets everything, n^j is two characters
        intervalData <- intervalData[nchar(intervalData$text) !=0, ] #absolutely gets everything, that is not a blank space
      } else { #specific categories specified
        intervalData <- intervalData[intervalData$text %in% categories, ]
      }

      fileNumber <- (intervalData$file)[1]

      df <- intervalData

      #attaching midpoint and plainTextname for later textgrid access
      df <- df %>% mutate(mid_point = (xmin + xmax)/2) #THIS IS FINE
      df <- df %>% mutate(plainTextName = plainTextname)

      textTiers <-  textGridDataFile[textGridDataFile$tier_type == "TextTier",]

      if (nrow(df) > 0){
        #1) using df min and max, isolate textTiers fragment
        #2) out of fragment, find textTier closest to midpoint
        frameNumberList <- list()
        for (midpoint in 1:length(df$mid_point)){
          min <- (df$xmin)[[midpoint]]
          max <- df$xmax[[midpoint]]
          textTierSection <- textTiers[textTiers$xmin >= min & textTiers$xmin <= max, ]

          if (nrow(textTierSection)!=0){
            frameNumber <- (textTierSection[which.min(abs(textTierSection$xmin - df$mid_point[[midpoint]])), ])$text
          }else{ #special case
            frameNumber <- (textTiers[which.min(abs(textTiers$xmin - df$mid_point[[midpoint]])), ])$text
          }
          frameNumberList = append(frameNumberList, frameNumber)
        }

        frameNumberList <- unlist(frameNumberList)
        df <- df %>% mutate(frame = frameNumberList)

        for (midpoint in 1:length(df$mid_point)){
          allRowsTextGrids <- rbind(allRowsTextGrids, data.frame(df[midpoint,])) #necessary in case multiple text grid files
        }
      }
    }else{
      #print("skipping this textgrid file")
    }

  }

  #extract xy data separately once we have all the data

  #allRowsTextGrids
  #textTiers
  if (length(allRowsTextGrids)==0){
    stop("the tiernames/categories that you specified either do not exist, or have zero traces")
  }  
  for(frame in 1:length(allRowsTextGrids$frame)){
    frameNumber = (allRowsTextGrids$frame)[[frame]]
    plainTextname <- allRowsTextGrids$plainTextName[[frame]]
    xyFileData <- (metaData$traces)$tongue$files[[plainTextname]][[frameNumber]]

    myFileAndFrameName <- paste(plainTextname, "_", frameNumber, sep = "")
    myVowelType <- (allRowsTextGrids$text)[[frame]]

    if (length(xyFileData)>0){
      for (mark in 1:length(xyFileData)){
        itemNumber <- mark
        xCoor <- xyFileData[[mark]]$x
        yCoor <- xyFileData[[mark]]$y

        #build data.frame
        appendedXYFrame <- c(myFileAndFrameName, itemNumber, myVowelType, yCoor, xCoor)
        myXY_data <- rbind(myXY_data, appendedXYFrame)
      }
    }
  }
  colnames(myXY_data) <- myXY_data[1, ] #replace auto generated heading
  myXY_data <- myXY_data[-1, ] #delete the heading that is in row 1
  myXY_data[ ,4] <- as.numeric(myXY_data[ ,4]) #x, y are ints for graphing
  myXY_data[ ,5] <- as.numeric(myXY_data[ ,5])
  
  #a[order(factor(a$x, levels = reference)),] #sorting given user input
  if (length(categories) != 0){
    myXY_data = myXY_data[order(factor(myXY_data$segment, levels = categories)), ]
  }
  
  return(myXY_data)
}
```

```{r}
#COLOR PALETTES:
#par(mfrow=c(1, 1))
#display.brewer.pal(9, "RdBu")
#brewer.pal(9, "RdYlBu")[[1]] #accessing individual colors
#https://www.datanovia.com/en/blog/the-a-z-of-rcolorbrewer-palette/
#display.brewer.all(colorblindFriendly = TRUE)
```
Sort in order by angle 
```{r}
#Return order of x values:
sortedOrder <- function(listX){ #Sort by angle
  order_x <- order(listX)
  return(order_x)
}
```

Calculate Intersection 
```{r}
calculateIntersection2 <- function(radiusPrevious, radiusNext, anglePrevious, angleRay, angleNext){

  differenceInTheta = angleNext - anglePrevious
  differenceInRay = angleRay - anglePrevious
  differenceInRadius = radiusNext - radiusPrevious
  
  return ( radiusPrevious + differenceInRay/differenceInTheta * differenceInRadius  )
}
```

Reading in data:
```{r}
# Import the text file
read_in_data <- function(extractedData){
  dataR <- extractedData
  
  #split by vowel type
  split_data <- split(dataR, dataR$segment)
  uniqueSegments <- unique(dataR$segment)
  
  dataOfEachCurveNNj <- list()
  for (segment in 1:length(uniqueSegments)){
    dataOfEachCurveNNj[[uniqueSegments[[segment]]]] <- split(split_data[[uniqueSegments[[segment]]]], split_data[[uniqueSegments[[segment]]]]$file_number)
  }
  
  return(dataOfEachCurveNNj)
}


get_unique_segments <- function(extractedData){
  dataR <- extractedData
  uniqueSegments <- unique(dataR$segment)

  return(uniqueSegments)
}
```

```{r}
identifyXAverage <- function(myXY_data){ #the x y axis are rotate left 90 degrees, so go based off x axis
  xmin =  min(myXY_data[c(5)])
  xmax =  max(myXY_data[c(5)])

  return((xmin + xmax)/2) 
}

identifyPlotBounds <- function(compiledList){
  xmin = 2
  xmax = -2
  ymin = 2
  ymax = -2
  
  for (segment in 1:length(compiledList)){
    for (trace in 1:length(compiledList[[segment]])){
      tempxmin <- min(compiledList[[segment]][[trace]][1,])
      tempxmax <- max(compiledList[[segment]][[trace]][1,])
      tempymin <- min(compiledList[[segment]][[trace]][2,])
      tempymax <- max(compiledList[[segment]][[trace]][2,])
      
      if (tempxmin < xmin){
        xmin <- tempxmin
      }
      
      if (tempxmax > xmax){
        xmax <- tempxmax
      }
      
      if (tempymin < ymin){
        ymin <- tempymin
      }
      
      if (tempymax > ymax){
        ymax <- tempymax
      }
    }
    
  }
  
  return(c(xmin, xmax, ymin, ymax))
}
#identifyPlotBounds(compiledList)
#use .5, use range, override (origin is -10) for shifting coordinates what about y but ok
```


Setting up 3D array for each segment: contains 1) x values 2) y values 3) angle in radians 4) radius
```{r}
formating_data <- function(dataOfEachCurveNNj, uniqueSegments, origin.x = .5, scaling.factor = 800/600){
  compiledList <- list()
  
  for (segment in 1:length(dataOfEachCurveNNj)){
    listofarrays <- list()
    for (i in 1:length(dataOfEachCurveNNj[[segment]])){ #for each curve
      n_rows <- 4 # no need for slope 
      n_cols <- length(dataOfEachCurveNNj[[segment]][[i]][,1])
      myCurveMatrix <- matrix(0, nrow = n_rows, ncol = n_cols) #each individual 2D array
      
      #adjusted x and y values around the center
      xvalues <- (dataOfEachCurveNNj[[segment]][[i]][c(5)][[1]] * scaling.factor) - (origin.x*scaling.factor) #include parameter for aspect ratio
      yvalues <- 1 - dataOfEachCurveNNj[[segment]][[i]][c(4)][[1]]
      
      
      anglevalues <- list()
      
      for (j in 1: n_cols){
        anglevalues <- append(anglevalues, atan2(yvalues[[j]], xvalues[[j]]) )
      }
      
      #sort by angle
      anglevalues <- unlist(anglevalues)
      
      theOrderForSorting <- sortedOrder(anglevalues)
      xvaluesSorted <- xvalues[theOrderForSorting]
      yvaluesSorted <- yvalues[theOrderForSorting]
      anglevalues <- anglevalues[theOrderForSorting]
      
      
      for (j in 1:n_cols){
          myCurveMatrix[1,j] <- xvaluesSorted[[j]]
          myCurveMatrix[2,j] <- yvaluesSorted[[j]]
          myCurveMatrix[3,j] <- anglevalues[[j]] #angle in relation to (0,0)
          myCurveMatrix[4,j] <- (yvaluesSorted[[j]]^2 + xvaluesSorted[[j]]^2)^.5 #radius in relation to (0,0)
          
      }
      
      listofarrays[[i]] <-myCurveMatrix
    }
    compiledList[[uniqueSegments[[segment]]]] <- listofarrays
  }
  return(compiledList)
}

```

create intersection matrix for each segment
where column is the angle, and the row is the value of the intersection of that angle for each trace in the segment, NA if none
```{r}
find_intersection_with_ray <- function(formatedData, dataOfEachCurveNNj, uniqueSegments, rayIncrement){ #ie compiledList
  
  matrixIntersection <- list()
  
  #start set up unique segment 
  for (segment in 1:length(uniqueSegments)){
    matrixIntersection[[uniqueSegments[[segment]]]] <- matrix(NA, nrow = length(formatedData[[segment]]), ncol = (3.14/rayIncrement))
  }
  #end set up
  
   #for each extending radius
    for (angleRay in seq(from=0, to= 3.14, by= rayIncrement)){
      
      for (segment in 1:length(uniqueSegments)){ #and for each segment
      
       for (individualTrace in seq(length(dataOfEachCurveNNj[[segment]]))){ #for each trace in the segment (alphabetical??)
         
         for (individualPoint in 1:(length(formatedData[[segment]][[individualTrace]][1, all()])-1)){ #for each point on the trace, find intersection
          #find the right two points the angle falls between for each individual trace
           
           if (angleRay <= formatedData[[segment]][[individualTrace]][3, individualPoint +1]){
             if (angleRay >= formatedData[[segment]][[individualTrace]][3, individualPoint]){ 
               
               #print("found intersection with the following ray angle")
               
               myIntersection <- calculateIntersection2(
                                                
                                                                    (formatedData[[segment]][[individualTrace]][4,individualPoint]), 
                                                                     (formatedData[[segment]][[individualTrace]][4,individualPoint + 1]), 
                                                                     (formatedData[[segment]][[individualTrace]][3,individualPoint]), 
                                                                     angleRay, 
                                                                     (formatedData[[segment]][[individualTrace]][3,individualPoint + 1]) 
                                                                     
                                                                     )
  
               #store value
               matrixIntersection[[segment]][[individualTrace, angleRay/rayIncrement]] <- myIntersection
               break
             }
           }
               
         }
         
     } #end individual trace for loop 
      
    } #end segment for loop
  
  } #end angle ray for loop 
  return(matrixIntersection) # columns for rays
} 
```

```{r}
#POST PROCESSING

my_palette1 <- c(
  "#FF0000", "#800080", "#0000FF",   # Red, Purple, Blue
  "#FF3333", "#993399", "#3366FF",   # Light Red, Light Purple, Light Blue
  "#FF6666", "#B366CC", "#6699FF",   # Medium Red, Medium Purple, Medium Blue
  "#FF9999", "#CC99FF", "#99BBFF",   # Dark Red, Dark Purple, Dark Blue
  "#FFCCCC", "#E6CCFF", "#CCE6FF"    # Lighter Red, Lighter Purple, Lighter Blue
)

plotStyleTraces <- function(matrixIntersection, compiledList, dataOfEachCurveNNj, uniqueSegments, palette = c(), rayIncrement, points.display = FALSE, mean.lines = TRUE, means.styles = c(), bands.fill = TRUE, bands.lines = FALSE, legend.position = "topleft", standard.deviation.styles = "l", pdf.filename = c(), png.filename = c(), plot.ticks = FALSE, plot.labels = FALSE, legend.size = 3, transparency = 0.37, bands.linewidth = 0.3, legend.linewidth = 5, means.linewidth = 3, tick.size = 2){
  
  plotbounds <- identifyPlotBounds(compiledList)
  
  standardDeviation <- list()
  averagedRX <- list()
  averagedRY <- list()
  
  for (segment in 1:length(uniqueSegments)){
    averagedRX[[uniqueSegments[[segment]]]] <- list()
    averagedRY[[uniqueSegments[[segment]]]] <- list()
    
    standardDeviation[[uniqueSegments[[segment]]]] <- list()
    
    for (ray in 1:length(matrixIntersection[[segment]][1, all()])){
      if (!is.na(colMeans(matrixIntersection[[segment]], na.rm = TRUE)[[ray]])){ #as long as there is a mean for each column, store the value
        averagedRX[[segment]] <- append(averagedRX[[segment]], cos(rayIncrement*ray) * colMeans(matrixIntersection[[segment]], na.rm = TRUE)[[ray]] )
        averagedRY[[segment]] <- append(averagedRY[[segment]], sin(rayIncrement*ray) * colMeans(matrixIntersection[[segment]], na.rm = TRUE)[[ray]] )
        standardDeviation[[segment]] <- append(standardDeviation[[segment]], sd(matrixIntersection[[segment]][all(), ray], na.rm = TRUE))
      } 
      else{
        averagedRX[[segment]] <- append(averagedRX[[segment]], NA)
        averagedRY[[segment]] <- append(averagedRY[[segment]], NA)
        standardDeviation[[segment]] <- append(standardDeviation[[segment]], NA)
      }
    }
    averagedRX[[segment]] <- unlist(averagedRX[[segment]])
    averagedRY[[segment]] <- unlist(averagedRY[[segment]])
    standardDeviation[[segment]] <- unlist(standardDeviation[[segment]])
  }

  
  #we now have the standard deviation
  xSDHigh <- list()
  xSDLow <- list()
  ySDHigh <- list()
  ySDLow <- list()  
  for (segment in 1:length(uniqueSegments)){
      
      xSDHigh[[uniqueSegments[[segment]]]] <- list()
      xSDLow[[uniqueSegments[[segment]]]] <- list()
      ySDHigh[[uniqueSegments[[segment]]]] <- list()
      ySDLow[[uniqueSegments[[segment]]]] <- list()
    
      for (i in 1:length(standardDeviation[[segment]])){
        
        if (!is.na(standardDeviation[[segment]][[i]])){
          xSDHigh[[segment]]<- append(xSDHigh[[segment]], cos(i*rayIncrement) *standardDeviation[[segment]][[i]] + averagedRX[[segment]][[i]])
          xSDLow[[segment]] <-  append( xSDLow[[segment]], -1*cos(i*rayIncrement) *standardDeviation[[segment]][[i]] + averagedRX[[segment]][[i]])
          
          ySDHigh[[segment]] <- append(ySDHigh[[segment]], sin(i*rayIncrement) *standardDeviation[[segment]][[i]] + averagedRY[[segment]][[i]])
          ySDLow[[segment]] <-  append(ySDLow[[segment]], -1*sin(i*rayIncrement) *standardDeviation[[segment]][[i]] + averagedRY[[segment]][[i]])
        
        }
        # else{
        #   xSDHigh[[segment]]<- append(xSDHigh[[segment]], NA)
        #   xSDLow[[segment]] <-  append( xSDLow[[segment]], NA)
        #   
        #   ySDHigh[[segment]] <- append(ySDHigh[[segment]], NA)
        #   ySDLow[[segment]] <-  append(ySDLow[[segment]], NA)
        # }
        
      }
  }

  xPlotAverage <- (plotbounds[[1]] + plotbounds[[2]])/2
  yPlotAverage <- (plotbounds[[3]] + plotbounds[[4]])/2
  
  x_ticks <- c(round(plotbounds[[1]],2), round(xPlotAverage,2), round(plotbounds[[2]],2))
  y_ticks <- c(round(plotbounds[[3]],2), round(yPlotAverage,2), round(plotbounds[[4]],2))
  #else {
    #still have "ticks" just way out of bounds so that they don't show up
    # x_ticks <- c(round(plotbounds[[1]],2) + 3, round(xPlotAverage,2) + 3, round(plotbounds[[2]],2) + 3)
    # y_ticks <- c(round(plotbounds[[3]],2) + 2, round(yPlotAverage,2) + 2, round(plotbounds[[4]],2) + 2)

  if(plot.labels == TRUE){
    x_ticks_lables <- c(round(plotbounds[[1]],2)*100, round(xPlotAverage,2)*100, round(plotbounds[[2]],2)*100)
    y_ticks_lables <- c(round(plotbounds[[3]],2)*100, round(yPlotAverage,2)*100, round(plotbounds[[4]],2)*100)
  }else{
    x_ticks_lables <- c(NA, NA, NA)
    y_ticks_lables <- c(NA, NA, NA)
  }
  
  if (length(pdf.filename)!=0){
     #pdf.options(encoding="MacRoman")
     #pdf.options(encoding = "utf-8")
     cairo_pdf(file = pdf.filename, family = "DejaVu Serif", width = 22 * (plotbounds[[2]] - plotbounds[[1]] - 0.05), height = 22 * (plotbounds[[4]] - plotbounds[[3]] -0.05))
    #cairo_pdf(file = pdf.filename, family = "DejaVu Serif")
    # cairo_pdf(file = pdf.filename, family = "DejaVu Serif", width = 22, height = 22)
  }
  
   if (length(png.filename)!=0){
      png(filename = png.filename, width = 4000*(plotbounds[[2]] - plotbounds[[1]] - 0.05), height = 4000 * (plotbounds[[4]] - plotbounds[[3]] -0.05), units = "px")
  }
 
  
  #par(pty = "s")
  par(mar = c(0, 4.4, 0, 1.25) + 2)
  
  
  plot(1, type = "n", xlab = "", ylab = "", ylim = c(plotbounds[[3]], plotbounds[[4]]), xlim = c(plotbounds[[1]], plotbounds[[2]]), xaxt = "n", yaxt = "n", asp = 1, family = "DejaVu Serif", cex.axis = 20)
  
  if(plot.ticks == TRUE){
    axis(1, at = x_ticks, labels = x_ticks_lables, cex.axis = tick.size)
    axis(2, at = y_ticks, labels = y_ticks_lables, cex.axis = tick.size)
  }else{
    axis(1, at = x_ticks, labels = x_ticks_lables, tck = 0, cex.axis = tick.size)
    axis(2, at = y_ticks, labels = y_ticks_lables, tck = 0, cex.axis = tick.size)
  }
  
  if (length(palette) == 0){
    numberOfColors <- length(uniqueSegments) + 2
    paletteColors <- (brewer.pal(numberOfColors, "Set1"))[2:numberOfColors] #PuRd nice
  
  }else{
    paletteColors <- palette
  }
  

  for (segment in 1:length(uniqueSegments)){
    #Shading
    x1 <- rev(unlist(lapply(xSDLow[[segment]], function(x) na.omit(x)), recursive = TRUE))
    y1 <- rev(unlist(lapply(ySDLow[[segment]], function(x) na.omit(x)), recursive = TRUE))


    x2 <- rev(unlist(lapply(xSDHigh[[segment]], function(x) na.omit(x)), recursive = TRUE))
    y2 <- rev(unlist(lapply(ySDHigh[[segment]], function(x) na.omit(x)), recursive = TRUE))


    #print(xSDHigh[[segment]])
    
    #Plot the upper lower standard devation lines
    if (bands.lines == TRUE){
      lines(x1, y1, type = standard.deviation.styles, col = paletteColors[[segment]], lwd = bands.linewidth)#, ylim = c(ymin, ymax), xlim = c(xmin, xmax))
      lines(x2, y2, type = standard.deviation.styles, col = paletteColors[[segment]], lwd = bands.linewidth)
    }
    
    if (bands.fill == TRUE){
      # Create a polygon to shade the region between the arches
      polygon(c(x1, rev(x2)), c(y1, rev(y2)), col = adjustcolor(paletteColors[[segment]], alpha.f = transparency), border = NA)
    }
    
    if (length(means.styles) < length(compiledList)){
      for (i in 1:(length(compiledList) - length(means.styles))){
        means.styles = append(means.styles, 1)
      }
    }

    if (mean.lines){
       if (!all(is.na(standardDeviation[[segment]]))){
        lines(averagedRX[[segment]], averagedRY[[segment]], type = "l", lty = means.styles[[segment]], col = paletteColors[[segment]] , lwd = means.linewidth) #you could also have black...
      }else{ #differentiating if there is a standard deviation band or not, currently no differation
        lines(averagedRX[[segment]], averagedRY[[segment]], type = "l", lty = means.styles[[segment]], col = paletteColors[[segment]] , lwd = means.linewidth)
      }
    }
    
    if (points.display){
      for (trace in 1:length(compiledList[[segment]])){
        points(compiledList[[segment]][[trace]][1, all()], compiledList[[segment]][[trace]][2, all()], type = "p", col = paletteColors[[segment]], asp = 1)
      }
    }
    
    # if (length(pdf.filename) == 0 && length(png.filename) == 0){
    #   legend.size = 0.6
    # }
    
    ltyNumerical = means.styles
    
    numberColumns = 1
    numberColumns = round(length(uniqueSegments)/5)
    if (numberColumns==0){
      numberColumns = 1
    }
    
    if (legend.position == "center"){
      # legend(xPlotAverage, yPlotAverage, legend = uniqueSegments,  col = paletteColors, cex = legend.size, bty = "n", lty=ltyNumerical, lwd = 5)
      legend("center", legend = uniqueSegments,  col = paletteColors, cex = legend.size, bty = "n", lty=ltyNumerical, lwd = legend.linewidth, ncol =  numberColumns)
    } else if (legend.position == "topleft"){
      # legend(plotbounds[[1]], plotbounds[[4]], legend = uniqueSegments, col = paletteColors, cex = legend.size, bty = "n", lty=ltyNumerical, lwd = 5)
      legend("topleft", legend = uniqueSegments, col = paletteColors, cex = legend.size, bty = "n", lty=ltyNumerical, lwd = legend.linewidth, ncol =  numberColumns)
    }else if (legend.position == "bottomright"){
      # legend((xPlotAverage + .5*plotbounds[[2]]), yPlotAverage,  legend = uniqueSegments, col = paletteColors, cex = legend.size, bty = "n", lty=ltyNumerical, lwd = 5, ncol = round(length(uniqueSegments)/5))
      legend("bottomright",  legend = uniqueSegments, col = paletteColors, cex = legend.size, bty = "n", lty=ltyNumerical, lwd = legend.linewidth, ncol =  numberColumns)
    }
  }
  
  if(length(pdf.filename)!=0 || length(png.filename)!=0 ){
    dev.off()
  }
  
}
```

PUTTING CONVERTING TO RADIANS ALL TOGETHER
```{r}
makeTracesPolar <- function(myXY_data, origin.algorithm = "BottomMiddle", origin.x = NA, scaling.factor = 800/600){
  
  uniqueSegments <- get_unique_segments(myXY_data)
  dataOfEachCurveNNj <- read_in_data(myXY_data)
  xaverage <- identifyXAverage(myXY_data)
  
  
  if (origin.algorithm == "BottomMiddle"){
    if (is.na(origin.x)){
      origin.x = .5
    }
  }else if (origin.algorithm == "BottomMean"){
    origin.x = xaverage
  }
  
  compiledList <- formating_data(dataOfEachCurveNNj, uniqueSegments, origin.x = origin.x, scaling.factor = scaling.factor)
  return(compiledList)
}

#polarTraces <- makeTracesPolar(myXY_data = myXY_data, origin.algorithm = "BottomMean")
#592, 412 : russian palatals
#800, 600 : others
```

```{r}
plotTraces <- function(myXY_data, compiledList, interval = 1, mean.lines = TRUE, points.display = FALSE, palette = c(), bands.lines = FALSE, bands.fill = TRUE, legend.position = "topleft", means.styles = c(), standard.deviation.styles = "l", plot.ticks = FALSE, plot.labels = FALSE, legend.size = 3, transparency = 0.37, pdf.filename = c(), bands.linewidth = 0.3, png.filename = c(), legend.linewidth = 5, means.linewidth = 3, tick.size = 2){
  
  rayIncrement = 3.14159/180 * interval
  
  uniqueSegments <- get_unique_segments(myXY_data)
  dataOfEachCurveNNj <- read_in_data(myXY_data)
  
  matrixIntersection <- find_intersection_with_ray(compiledList, dataOfEachCurveNNj, uniqueSegments, rayIncrement)
    
  plotStyleTraces(matrixIntersection = matrixIntersection, compiledList = compiledList, dataOfEachCurve = dataOfEachCurveNNj, uniqueSegments = uniqueSegments, rayIncrement = rayIncrement, mean.lines = mean.lines, points.display = points.display, palette = palette, bands.lines = bands.lines, legend.position = legend.position, bands.fill = bands.fill, means.styles = means.styles, standard.deviation.styles = standard.deviation.styles, plot.ticks = plot.ticks, legend.size = legend.size, transparency = transparency, pdf.filename = pdf.filename, bands.linewidth = bands.linewidth, plot.labels = plot.labels, png.filename = png.filename, legend.linewidth = legend.linewidth, means.linewidth = means.linewidth, tick.size = tick.size)
}
```

```{r}
#directory_name <- "/Users/yayaorigami/Documents/plinguistics/IETestData"
#directory_name <- "/Users/yayaorigami/Documents/plinguistics/ultrasound-data-RussianPalatals-202012/"
categories <- c("o","y", "e", "i")
categories <- c('u', 'ɯ', 'o', 'ɑ', 'i', 'y', 'e', 'ø')
#categories <- c("nʲ", "n")

myXY_data <- loadTraces(directory_name = "/Users/yayaorigami/Documents/plinguistics/updatedRussianPalatals/ultrasound-data-RussianPalatals-202012/", "cons", c("n", "l"))
myXY_data
```
```{r}
polarTraces <- makeTracesPolar(myXY_data = myXY_data, origin.algorithm = "BottomMean", scaling.factor = 592/412)
plotTraces(myXY_data = myXY_data, compiledList = polarTraces, points.display = FALSE, bands.lines = TRUE, bands.fill = TRUE, transparency = 0.37, mean.lines = TRUE, means.styles =  c(2, 1, 2, 1, 1, 2, 1, 2), legend.position = "topleft", plot.labels = TRUE, legend.size = 1.2, legend.linewidth = 1.5, means.linewidth = 1.5, tick.size = 1.2, bands.linewidth = .3)#, png.filename = "ok4.png")
```
Set1, png resolution, topleft, tiername bug

Checking sizes IS A FUNCTION (that you will not be able to currently use)
```{r}
#dataOfEachCurveNNj <- read_in_data_file_path(file_path)

dataOfEachCurveNNj <- read_in_data(myXY_data)
uniqueSegments <- get_unique_segments(myXY_data)

checkSizes <- function(dataOfEachCurveNNj, uniqueSegments){
  for (segment in 1:length(dataOfEachCurveNNj)){
    print(names(dataOfEachCurveNNj)[[segment]])
    print(length(dataOfEachCurveNNj[[segment]]))
    sizes <- list()
    print("number of points per curve")
    for (curve in 1:length(dataOfEachCurveNNj[[segment]])){
      sizes <- append(sizes, length(dataOfEachCurveNNj[[segment]][[curve]][[1]]) )
    }
    print(unlist(sizes))
  }
}

checkSizes(dataOfEachCurveNNj, uniqueSegments)
uniqueSegments
```
Testing labelling function:
```{r}
split_data <- split(myXY_data, myXY_data$segment)
uniqueSegments <- unique(myXY_data$segment)
#split_data
#uniqueSegments

uniqueSegments[[2]]
split_data[[uniqueSegments[[2]]]]


dataOfEachCurveNNj <- list()
for (segment in 1:length(uniqueSegments)){
  dataOfEachCurveNNj[[uniqueSegments[[segment]]]] <- split(split_data[[uniqueSegments[[segment]]]], split_data[[uniqueSegments[[segment]]]]$file_number)
}

dataOfEachCurveNNj[[2]]
```

